Analytic Occultation Light Curves in Reflected Light

### 2020-04-22 tex/ms.tex
better normalization for lambertian spheres

%
\section{Introduction}
\label{sec:intro}
%
This is an extension of \citet{Luger2019}.

@@

@@

@@

\subsection{Review of the \starry algorithm in emitted light}
\label{sec:starry-review}
%
Without loss of generality, assume the body whose flux we wish to compute
has radius unity and sits at the origin of a right-handed Cartesian coordinate
system in some frame $\mathcal{F}_0$. In this frame,
the surface (emitted) intensity field of the body is described by a
vector $\mathbf{y}$ of coefficients in the spherical harmonic
basis $\by$:
%
\begin{align}
    \label{eq:by}
    \by & =
    \begin{pmatrix}
        Y_{0, 0}  &
        Y_{1, -1} & Y_{1, 0}  & Y_{1, 1} &
        Y_{2, -2} & Y_{2, -1} & Y_{2, 0} & Y_{2, 1} & Y_{2, 2} &
        \cdot\cdot\cdot
    \end{pmatrix}^\top
    \quad,
\end{align}
%
where the component at index $n$ is $Y_{l,m}$ with
%
\begin{align}
    \label{eq:l-m}
    l & = \floor*{\sqrt{n}}
    \nonumber               \\
    m & = n - l^2 - l
    \quad.
\end{align}
%
An observer views the body from a large distance in the sky frame
$\mathcal{F}$, in which the $x$-axis points to the right,
the $y$-axis points up, and the $z$-axis points out of the sky
toward the observer. Following \citet{Luger2019}, if
an occultor of radius $r_o$ is located at sky position $(x_o, y_o)$,
we compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \mathbf{A} \mathbf{R}' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, from right to left, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$
given the body's inclination $\text{I}$, obliquity
$\Lambda$, and rotational phase $\Theta$
\citep[Appendix C in][]{Luger2019},
%
$\mathbf{R}' = \mathbf{R}'(x_o, y_o)$ rotates the body on the plane
of the sky into the integration frame $\mathcal{F}'$, in which the
occultor lies along the $+y'$-axis,
%
$\mathbf{A}$
\citep[Equation~B13 in][]{Luger2019}
is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed,
whose component at index $n$ is
%
\begin{align}
    \label{eq:bg}
    \tilde{g}_{n}(x, y) & =
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \text{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \, \text{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
with
%
\begin{align}
    \label{eq:mu-nu}
    \mu & \equiv l - m
    \nonumber          \\
    \nu & \equiv l + m
\end{align}
%
and
%
\begin{align}
    \label{eq:z}
    z(x, y) \equiv \sqrt{1 - x^2 - y^2}
    \quad,
\end{align}
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
\citep[Equation~26 in][]{Luger2019}, with $b_o = \sqrt{x_o^2 + y_o^2}$.

@@

\subsection{Adapting the algorithm to the reflected light case}
\label{sec:adapting-starry}
%
In order to compute light curves in reflected light, we must make two
modifications to the \starry algorithm. First,
the expressions above assume that the coefficient vector
$\mathbf{y}$ describes the \emph{emissivity} of the body, which (in the
absence of limb darkening) is assumed to be Lambertian, i.e., all points on the
surface emit equally in all directions.
Here, we wish to derive the solution for the flux in the case of Lambertian
reflectance, in which case the vector $\mathbf{y}$ is taken to describe the
surface \emph{reflectivity} (or \emph{Bond albedo}).

Second, we must explicitly model the illumination of the body. We assume the
body is illuminated by a point-like source of unit luminosity, in which case
the radiance at any
point on the surface is proportional to the cosine of the angle between
the incident light and the surface normal. Points for which
this angle is greater than $\nicefrac{\pi}{2}$ are unilluminated and
therefore have radiance of zero.
%
If the point-like illumination source is placed at sky coordinates
$(x_s, y_s, z_s)$, the day/night terminator on the body is a half-ellipse
of semi-major axis unity that is fully described by its (signed) semi-minor
axis,
%
\begin{proof}{}
    \label{eq:b}
    b = -\frac{z_s}{r_s}
    \quad,
\end{proof}
%
where $r_s = \sqrt{x_s^2 + y_s^2 + z_s^2}$ is the distance to the source,
%
and the angle $\theta$ by which its semi-major axis is rotated away from the
$+x$-axis (a frame-dependent quantity, defined below).
Given this formulation, and assuming that
$r_s \gg 1$,
it is straightforward to show that the illumination
$\mathcal{I}$ at a point $(x, y)$ on the projected disk of the body is given
by the function
%
\begin{proof}{}
    \label{eq:illum}
    \mathcal{I}(b, \theta, r_s; x, y)&=
    \text{max}\bigg( 0, I(b, \theta, r_s; x, y) \bigg)
\end{proof}
%
where
%
\begin{proof}{}
    \label{eq:illum_poly}
    I(b, \theta, r_s; x, y) &= \frac{1}{4\pi r_s^2}
    \bigg(
    -b_c\sin\theta x + b_c\cos\theta y - bz(x, y)
    \bigg)
\end{proof}
%
with $b_c \equiv \sqrt{1 - b^2}$ and $z(x, y) = \sqrt{1 - x^2 - y^2}$.

@@

    \frac{1}{4\pi r_s^2}

>>>>

%
\section{Introduction}
\label{sec:intro}
%
Introduction here.

@@

Our goal in this paper is to derive analytic expressions for the
flux received by a distant observer from a non-uniform Lambert sphere
illuminated by a monochromatic source that may or may not be
occulted by a (possibly different) spherical body. This applies, for example,
to the
case of exoplanet phase curves, exoplanet secondary
eclipse (occultation) light curves, or exoplanet-moon and exoplanet-exoplanet
occultations, all seen in reflected light. We explicitly assume the reflecting
body is Lambertian, i.e., it scatters light isotropically, although with an
intensity that varies across the surface according to a spatially-dependent
albedo $A$. Throughout this paper, we will take $A$ to mean the \emph{spherical}
albedo, the fraction of power incident on a body at a given wavelength
that is scattered back out to space (in all directions).
Note that the spherical albedo
is closely related to the Bond albedo: the Bond albedo is the stellar
flux-weighted integral of $A(\lambda)$ over all wavelengths $\lambda$. For
more details, see \S4.5 of \citet{Schwartz2015}.

As in \citet{Luger2019}, we compute fluxes by first expanding the surface
in terms of spherical harmonics.
While in \citet{Luger2019} we expanded the emissivity of the surface, here
we instead expand the spherical albedo $A$.
Specifically, if $\bvec{y}$ is the vector
of spherical harmonic coefficients describing the albedo anywhere on
the surface and
$\by$ is the spherical harmonics basis (Equation~\ref{eq:by}),
the albedo $A$ at a point $(x, y)$ on the sky-projected disk of the body
is given by the dot product
%
\begin{align}
    \label{eq:albedo}
    A(x, y) = \by^\top (x, y) \, \bvec{y}
    \quad.
\end{align}
%
The flux measured from this body is the surface integral over the visible
portion of the body's projected disk of the albedo $A$
times the illumination profile $\mathcal{I}$ of the surface, given by
Lambert's law as
%
\begin{align}
    \label{eq:LambertsLaw}
    \mathcal{I}(\phi) = \mathcal{I}_0 \, \text{max}\big( 0, \cos\phi \big)
    \quad,
\end{align}
%
where $\mathcal{I}_0$ is the peak illumination (at the sub-source point),
which we discuss in more detail in \S\ref{sec:adapting-starry}.

@@

\section{Extended illumination source}
\label{sec:extended}
Planets to consider: 55 Cnc e, Kelt-9b.
The angular extent of the terminator past $\nicefrac{\pi}{2}$ is
%
\begin{align}
    \tau = \arcsin\left( \frac{\Delta R}{d} \right)
\end{align}

\section{Non-Lambertian scatterers}
\label{sec:nonlambertian}
In principle, we can just change $\mathbf{I}$.

@@

\subsection{Review of the \starry algorithm in emitted light}
\label{sec:starry-review}
%
Without loss of generality, assume the body whose flux we wish to compute
has radius unity and sits at the origin of a right-handed Cartesian coordinate
system in some frame $\mathcal{F}_0$. In this frame,
the surface (emitted) intensity field of the body is described by a
vector $\mathbf{y}$ of coefficients in the spherical harmonic
basis $\by$:
%
\begin{align}
    \label{eq:by}
    \by(x, y) & =
    \begin{pmatrix}
        Y_{0, 0}  &
        Y_{1, -1} & Y_{1, 0}  & Y_{1, 1} &
        Y_{2, -2} & Y_{2, -1} & Y_{2, 0} & Y_{2, 1} & Y_{2, 2} &
        \cdot\cdot\cdot
    \end{pmatrix}^\top
    \quad,
\end{align}
%
where the component at index $n$ is the spherical harmonic $Y_{l,m}(x, y)$ with
%
\begin{align}
    \label{eq:l-m}
    l & = \floor*{\sqrt{n}}
    \nonumber               \\
    m & = n - l^2 - l
    \quad.
\end{align}
%
The spherical harmonics are traditionally expressed in spherical coordinates,
but for our purposes
it is more conventient to express them in Cartesian coordinates on
the sky-projected disk, in which case they are simply polynomials in
$x$, $y$, and $z$ \citep[see Appendix~A in][]{Luger2019}.

An observer views the body from a large distance in the sky frame
$\mathcal{F}$, in which the $x$-axis points to the right,
the $y$-axis points up, and the $z$-axis points out of the sky
toward the observer. Following \citet{Luger2019}, if
an occultor of radius $r_o$ is located at sky position $(x_o, y_o)$,
we compute the visible flux $f$ from
%
\begin{align}
    \label{eq:sTARRy}
    f = \sTe \mathbf{A} \mathbf{R}' \mathbf{R} \mathbf{y}
    \quad,
\end{align}
%
where, from right to left, $\mathbf{R} = \mathbf{R}(\text{I}, \Lambda, \Theta)$
is a Wigner rotation matrix that rotates $\bvec{y}$ from $\mathcal{F}_0$
to the sky frame $\mathcal{F}$
given the body's inclination $\text{I}$, obliquity
$\Lambda$, and rotational phase $\Theta$
\citep[Appendix C in][]{Luger2019},
%
$\mathbf{R}' = \mathbf{R}'(x_o, y_o)$ rotates the body on the plane
of the sky into the integration frame $\mathcal{F}'$, in which the
occultor lies along the $+y'$-axis,
%
$\mathbf{A}$
\citep[Equation~B13 in][]{Luger2019}
is the change-of-basis matrix from $\by$
to the \emph{Green's basis} $\bg$ in which the integrals are computed,
whose component at index $n$ is
%
\begin{align}
    \label{eq:bg}
    \tilde{g}_{n}(x, y) & =
    \begin{dcases}
        %
        \frac{\mu+2}{2}x^\frac{\mu}{2} y^\frac{\nu}{2}
         & \qquad \mu, \nu \, \text{even}
        \\[1em]
        %
        z(x, y)
         & \qquad \mu = \nu = 1
        \\[1em]
        %
        3x^{l-2}yz(x, y)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \frac{\mu + \nu}{2} \, \text{even}
        \\[1em]
        %
        z(x, y)
        \bigg(
        -x^{l-3} + x^{l-1} + 4x^{l-3}y^2
        \bigg)
         & \qquad \nu \, \text{odd}, \,
        \mu = 1, \,
        \, \text{odd}
        \\[1em]
        %
        z(x, y)
        \bigg(
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu-1}{2}
        \ - \
        \frac{\mu-3}{2} x^\frac{\mu-5}{2} y^\frac{\nu+3}{2}
        \\
        \qquad\qquad \ - \
        \frac{\mu+3}{2} x^\frac{\mu-1}{2} y^\frac{\nu-1}{2}
        \bigg)
         & \qquad \text{otherwise}
        \quad,
    \end{dcases}
\end{align}
%
with
%
\begin{align}
    \label{eq:mu-nu}
    \mu & \equiv l - m
    \nonumber          \\
    \nu & \equiv l + m
\end{align}
%
and
%
\begin{align}
    \label{eq:z}
    z(x, y) \equiv \sqrt{1 - x^2 - y^2}
    \quad,
\end{align}
%
and $\sTe = \sTe(b_o, r_o)$ is the vector of solutions to the integral over
the projected visible disk of the body for each term in $\bg$
\citep[Equation~26 in][]{Luger2019}, with $b_o = \sqrt{x_o^2 + y_o^2}$.

@@

\subsection{Adapting the algorithm to the reflected light case}
\label{sec:adapting-starry}
%
In order to compute light curves in reflected light, we must make two
modifications to the \starry algorithm. First,
the expressions above assume that the coefficient vector
$\mathbf{y}$ describes the \emph{emissivity} of the body, which (in the
absence of limb darkening) is assumed to be Lambertian, i.e., all points on the
surface emit equally in all directions.
Here, we wish to derive the solution for the flux in the case of Lambertian
reflectance, in which case the vector $\mathbf{y}$ is taken to describe the
spherical albedo of the surface, $A$.

Second, we must explicitly model the illumination of the body. We assume the
body is illuminated by a point-like source whose flux measured by the observer
is unity. In this case, the observed intensity at any
point on the surface is proportional to the cosine of the angle $\phi$ between
the incident light and the surface normal. Points for which
$\phi \ge \nicefrac{\pi}{2}$ are unilluminated and
therefore have an intensity of zero.
%
If the point-like illumination source is placed at sky coordinates
$(x_s, y_s, z_s)$ in units of the radius of the illuminated body,
the day/night terminator on the body is a half-ellipse
of semi-major axis unity that is fully described by its (signed) semi-minor
axis,
%
\begin{proof}{}
    \label{eq:b}
    b = -\frac{z_s}{r_s}
    \quad,
\end{proof}
%
where $r_s = \sqrt{x_s^2 + y_s^2 + z_s^2}$ is the distance to the source,
%
and the angle $\theta$ by which its semi-major axis is rotated away from the
$+x$-axis (a frame-dependent quantity, defined below).
Given this formulation, and assuming that
$r_s \gg 1$,
it is straightforward to show that the illumination
$\mathcal{I}$ at a point $(x, y)$ on the projected disk of the body is given
by the function
%
\begin{proof}{}
    \label{eq:illum}
    \mathcal{I}(b, \theta, r_s; x, y)&=
    \text{max}\bigg( 0, I(b, \theta, r_s; x, y) \bigg)
\end{proof}
%
where
%
\begin{proof}{}
    \label{eq:illum_poly}
    I(b, \theta, r_s; x, y) &= \frac{1}{\pi r_s^2}
    \bigg(
    -b_c\sin\theta x + b_c\cos\theta y - bz(x, y)
    \bigg)
\end{proof}
%
with $b_c \equiv \sqrt{1 - b^2}$ and $z(x, y) = \sqrt{1 - x^2 - y^2}$.
A quick note on the normalization of Equation~(\ref{eq:illum_poly}):
if we place the illumination source along the $+z$-axis at $(0, 0, z_s)$,
the body is seen at full phase, so $b = -1$, $b_c = 0$, and
%
\begin{proof}{}
    \mathcal{I}_\text{full}(x, y) = \frac{\sqrt{1 - x^2 - y^2}}{\pi z_s^2}
    \quad.
\end{proof}
%
Integrating this over the full disk, we obtain the reflected flux measured
by the observer if the body has unit spherical albedo,
%
\begin{proof}{}
    \mathcal{f}_\text{full} &=
    \int_{-1}^{1}
    \int_{-\sqrt{1 - x^2}}^{\sqrt{1 - x^2}}
    \frac{\sqrt{1 - x^2 - y^2}}{\pi z_s^2}
    \,
    \dd y
    \,
    \dd x
    \nonumber \\[0.5em]
    &= \left(\frac{2}{3}\right) \frac{1}{z_s^2}
    \quad,
\end{proof}
%
recalling that $z_s$ is measured in units of the illuminated body's
radius. The factor of $\nicefrac{2}{3}$ is
precisely the ratio of the geometric albedo to the spherical
albedo of a Lambert sphere \citep[see, e.g.][]{Seager2010}. If we placed
the source at $z_s = 1$, reflection off the illuminated body would increase
the total flux measured at the observer by $\nicefrac{2}{3}$ (since, by
assumption, the source flux at the observer is unity). As a sanity check,
if all radiation incident on the body were instead scattered back toward the
source (and hence toward the observer), $f_\text{full}$ would instead be unity:
the flux at the observer would \emph{double}.
\xxx{Is this explanation satisfactory? Does it make sense?}

@@

    \frac{1}{\pi r_s^2}
